---
  title: "Build EAP Plan Item Analysis Data"
author: "Analytics Unit"
date: "`r format(Sys.Date(), '%B %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---
  
  ```{r setup, cache=TRUE, results='hide', message=FALSE, warning=FALSE}

knitr::opts_chunk$set(echo = TRUE)
options(scipen = 999)

# Load libraries

require(odbc)
require(tidyverse)
require(dbplyr)
require(here)
require(rstudioapi)
require(janitor)
require(lubridate)
require(writexl)


## pull data

raw_plans <- read_rds("O:/Analytics Unit/Projects/EST/Employment Action Plan/data/raw_plans.rds")


# Check the referral program field to see how many times CORP comes up

raw_plans %>%
  count(REFERRED_PROGRAM)


# Data wrangling:
# Clean column headers and date fields
# Create start and end week aggregations
# Create a TRUE/FALSE flag to identify custom basic plan items

plans <- raw_plans %>%
  clean_names() %>%
  mutate_if(is.POSIXct, as_date) %>%
  mutate(
    referred_program = case_when(is.na(referred_program) ~ "MLTSD",
                                 TRUE ~ referred_program),
    plan_start_week = floor_date(actualstartdate, "1 week"),
    plan_end_week = floor_date(actualenddate, "1 week"),
    plan_duration = (actualenddate - actualstartdate) + 1,
    custom_plan = case_when(plan_item_name == "Custom Basic Plan Item" ~ TRUE,
                            TRUE ~ FALSE))

# Export at the plan item level

saveRDS(plans, here("data/EAP_plan_items.rds"))

```


Create a dataset for export that aggregates plan item data to the EAP level.

* plan_num: total number of plan items
* plan_total_cost: total cost of all plan items
* plan_total_duration: total duration of all plan items combined
* sub_...: number of plan items associated with each sub-goal

```{r}

# Make a temporary dataset of the number of plans per sub-goal for each EAP

temp_subgoals <- plans %>%
  count(casereference, sub_goal) %>%
  pivot_wider(names_from = "sub_goal",
              values_from = "n") %>%
  rename(sub_EAS = `Employment Assistance Services`,
         sub_JFS = `Jobseeker Financial Supports`,
         sub_RET = Retention,
         sub_SDM = `Skills Development - Ministry Delivered Programs`,
         sub_SDO = `Skills Development - Other`,
         sub_SPS = `Specialized Services`,
         sub_LST = `Life Stabilization`,
         sub_EFS = `Employer Financial Supports`)


# Make EAP-level dataset of plan item data for each EAP:

EAP_plans_export <- plans %>%
  group_by(casereference) %>%
  summarise(plan_num = n(),
            plan_total_cost = sum(actualcost),
            plan_total_duration = sum(plan_duration)) %>%
  left_join(temp_subgoals, by = "casereference")


# Export

saveRDS(EAP_plans_export, here("data/EAP_plans.rds"))

```





####### my analysis

plans %>%
  filter(!is.na(plan_item_name)) %>%
  group_by(plan_item_name) %>%
  summarize(cnt = n()) %>%
  mutate(pct = cnt/sum(cnt)) %>%
  arrange(desc(pct))






plans %>%
  count(plan_item_name) %>%
  adorn_totals %>%
  arrange(desc(n))
```{r}

####examine plan item costs by stream

test <- merge(plans, select(eaps,c("casereference","stream")), by = "casereference")




test %>%
            filter(actualcost>0) %>%
  group_by(sub_goal,plan_item_name,stream) %>%
  summarize(total_cost = sum(actualcost)) %>%
  ggplot(aes(reorder(plan_item_name,-total_cost),total_cost))+geom_col()+
  theme(axis.text.x = element_text(angle=90, hjust=1))+facet_wrap(~stream)



test %>%
  filter(actualcost>0) %>%
  filter(actualcost<20000) %>%
  group_by(stream,sub_goal) %>%
  summarize(total_cost = sum(actualcost)) %>%
  mutate(pct = total_cost/sum(total_cost)) %>%
  ggplot(aes(reorder(stream,-pct),pct,fill=sub_goal))+geom_bar(position = "stack",stat="identity")
  
 
  
  



```
  
  
  
  
  
  